---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r setoptions, echo=FALSE}
options(scipen = 1, digits = 2)
```


## Loading and preprocessing the data
The data was taken from the forked repository.
```{r}
unzip("activity.zip")
activity <- read.csv("activity.csv", as.is=TRUE)
activity$date <- as.Date(activity$date)
```



## What is mean total number of steps taken per day?
```{r}
totalStepsPerDay <- aggregate(activity$steps ~ activity$date, FUN=sum)
colnames(totalStepsPerDay) = c("Date", "Steps")
xlabel <- "Total number of steps taken each day"
hist(totalStepsPerDay$Steps, main = NA, xlab = xlabel)

meanTotalStepsPerDay <- mean(totalStepsPerDay$Steps)
medianTotalStepsPerDay <- median(totalStepsPerDay$Steps)
```
The mean of total number of steps taken per day is: `r meanTotalStepsPerDay`

The median of total number of steps taken per day is: `r medianTotalStepsPerDay`




## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
